Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-07-29 Final_Manuscript.Rmd
Added .bib citations to #Limitations and Future Directions. Fixed CITE, CITE, CITE's

There is evidence that the experience of state self-objectification in mixed-gender contexts (stranger and romantic) and within the self in any moment (trying on a swimsuit) has negative consequences, but what about in the context of interaction with other women? Is the self-objectification experience in mixed-gender interactions the same as that experienced in same-gender interactions? The ample research demonstrating that the male gaze has a particularly detrimental effect would suggest no [@saguyetal2010, @calogero2004test; @gervais2011you; @gay2010my]. Women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], but when they do it, does it lead to self-objectification in the same way men's objectification of women does [@garcia2016objectification]? If so, does the self-objectification experienced in these same-gender interactions have the same negative consequences for authenticity in that interaction?

@@

As is the case in all correlational studies, we cannot be sure in the causal direction. It could be that women's SSO causes them to be objectified by their interaction partner. This interpretation holds up given that, in both the @garcia2016objectification study and in the current study, it is *women's* SSO this relates to being objectified by one's partner. However, Objectification Thoery [@robertsfredrickson] as past experimental studies [@saguyetal2010; CITE] suggest that the causal direction is indeed in the direction whereby other-objectification causes SSO. Although, importantly, the current study found evidence for the partner effect of other-objectification on SSO, there was a lack of evidence for the connections between SSO and downstream negative consequences.

Where the results of the current study diverge most notably from the results of studies testing interpersonal objectification among mixed-gender dyads is the lack of evidence for relationships between SSO and felt inauthenticity. In addition, we found no evidence of *direct* or *indirect* relationships between SSO and cognitive functioning, relationship agency, and career aspirations. Nor was there evidence of *direct* or *indirect* relationships between other-objectification and any of these outcomes. This is somewhat suprising given the plentiful evidence linking SSO and cognitive functioning [@fredrickson1998swimsuit; @quinn2011performance] and the small, but extant research on interpersonal other-ojectification and cognitive functioning [@logel2009interacting; @garcia2016objectification]. This lack of evidence could potentially signal diverging processs between women's experiences with interpersonal objectification from men and interpersonal objectification from women. There is quite a bit of evidence suggesting that the male gaze is particularly detrimental [CITE; CITE; CITE], and perhaps the self-objectification experienced within an interaction with a women is not as detrimental as the self-objectificatoin experienced within an interaction with a man. We need to be careful not to interpret a null result as evidence of no relationship.

@@

LANEE PLEASE ADD THESE AS BIBTEX CITATIONS

Previous research has found that when compared to heterosexual women, lesbian women report less concern with physical appearance [Siever, 1994; Strong, Williamson, Netemeyer, & Geer, 2000], less internalization of sociocultural beauty norms [Bergeron & Senn, 1998; Share & Mintz, 2002], and less self-objectification [Brownlow, 1997; Noffsinger-Frazier, 2004]. None of these studies examined the relationship between self-objectification and experiences of cultural sexual objectification. Thus, it is unclear whether lesbians indeed experience similar levels of cultural sexual objectification but internalize them less than heterosexual women do. Based on theoretical literature suggesting that lesbians may internalize cultural sexual objectification less than do heterosexual women [e.g., Brown, 1987; Dworkin, 1988; Pitman, 1999; Rothblum, 1994; Siever, 1994; Striegel-Moore et al., 1986]

However, since both men and women are socialized in a culture that sexually objectifies women, both men and women may come to internalize this socialization and sexually objectify women. Women come to experience not only themselves, but other women as well, through the “eyes of the indeterminate observer” [Kaschak, 1992, p. 178]. Indeed, recent research has found that the more women self- objectify, the more they objectify other women [CITE], although not to the degree exhibited by men; that is, men were found to objectify women significantly more than women objectify other women [Strelan & Hargreaves, 2005].

>>>>

There is evidence that the experience of state self-objectification in mixed-gender contexts (stranger and romantic) and within the self in any moment (trying on a swimsuit) has negative consequences, but what about in the context of interaction with other women? Is the self-objectification experience in mixed-gender interactions the same as that experienced in same-gender interactions? The ample research demonstrating that the male gaze has a particularly detrimental effect would suggest no [@saguyetal2010, @calogero2004test; @gervais2011you; @gay2010my; @yilmaz2019whose]. Women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], but when they do it, does it lead to self-objectification in the same way men's objectification of women does [@garcia2016objectification]? If so, does the self-objectification experienced in these same-gender interactions have the same negative consequences for authenticity in that interaction?

@@

As is the case in all correlational studies, we cannot be sure in the causal direction. It could be that women's SSO causes them to be objectified by their interaction partner. This interpretation holds up given that, in both the @garcia2016objectification study and in the current study, it is *women's* SSO this relates to being objectified by one's partner. However, Objectification Theory [@robertsfredrickson] as past experimental studies [@saguyetal2010; CITE] suggest that the causal direction is indeed in the direction whereby other-objectification causes SSO. Although, importantly, the current study found evidence for the partner effect of other-objectification on SSO, there was a lack of evidence for the connections between SSO and downstream negative consequences.

Where the results of the current study diverge most notably from the results of studies testing interpersonal objectification among mixed-gender dyads is the lack of evidence for relationships between SSO and felt inauthenticity. In addition, we found no evidence of *direct* or *indirect* relationships between SSO and cognitive functioning, relationship agency, and career aspirations. Nor was there evidence of *direct* or *indirect* relationships between other-objectification and any of these outcomes. This is somewhat suprising given the plentiful evidence linking SSO and cognitive functioning [@fredrickson1998swimsuit; @quinn2011performance] and the small, but extant research on interpersonal other-ojectification and cognitive functioning [@logel2009interacting; @garcia2016objectification]. This lack of evidence could potentially signal diverging processs between women's experiences with interpersonal objectification from men and interpersonal objectification from women. There is quite a bit of evidence suggesting that the male gaze is particularly detrimental [@calogero2004test; @robertsfredrickson; @gervais2011you; @gervais2013my], and perhaps the self-objectification experienced within an interaction with a women is not as detrimental as the self-objectificatoin experienced within an interaction with a man. We need to be careful not to interpret a null result as evidence of no relationship.

@@

Previous research has found that when compared to heterosexual women, lesbian women report less concern with physical appearance [@siever1994sexual; @strong2000eating], and less self-objectification [@brownlow1998relationship; @noffsinger2004objectification]. However none of these studies examined the relationship between self-objectification and experiences of sexual objectification. Thus, it is unclear whether lesbians indeed experience similar levels of cultural sexual objectification but internalize them less than heterosexual women do. Consistent with previous reseach @hill2008examining determined lesbians existed less physical appearance concerns compared to heterosexual women, however there was no difference found that lesbian women self-objectify less than heterosexual women and they did not find that sexual orientation moderates the relationshiotn between sexual objecitifcantion and self-objectification This contradicts older theoretical literature that suggests that lesbians internalize cultural sexual objectification less than do heterosexual women [@brown1987lesbians; @pitman1999body; @lum1994lesbians; @siever1994sexual]

However, since both men and women are socialized in a culture that sexually objectifies women, both men and women may come to internalize this socialization and sexually objectify women. Indeed, recent research has found that the more women self-objectify, the more they objectify other women [@loughnan2015exploring; @puvia2013being], although not to the degree exhibited by men; that is, men were found to objectify women significantly more than women objectify other women [@strelan2005women].
